---
title: "xmlGenerator_basics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xmlGenerator_basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

`xmlGenerator` is an R package intended to transform one set of Microsoft Forms data into xml that is then imported into an EndNote library. `xmlGenerator` only works with one data structure that has a specific set of columns and column names and is not meant to be a a generic data frame to xml converter. If you need tools to create xml from data frames *in general*, `library(xml2)` is a good place to look.  

# Getting started  

`xmlGenerator` is available in [this GitHub repo](https://github.com/cw-data/xmlGenerator) and, as of 2023-02-20, is not published on CRAN.
Use `devtools` to download `xmlGenerator`.

```{r setup}
#----- install package dependencies:
# install.packages("devtools")
# or
# renv::install("devtools")
library(devtools)
library(stringr)
library(xml2)
library(dplyr)
library(data.table)
library(readxl)

#----- install xmlGenerator from github:
devtools::install_github("https://github.com/cw-data/xmlGenerator")
library(xmlGenerator)
```

# A minimal reproducible workflow  

An example dataset is included with `xmlGenerator`. This example dataset allows a user to:  
1) see the required format of their input data (`View(example_data)`)
2) see the outputs of `xmlGenerator` functions ((`View(forms_spreadsheet)`, `View(record_list)`, `cat(as.character(xml_output))`)
```{r}
#--------------------------------
# A minimal reproducible workflow that's included with library(xmlGenerator)
# Purpose: include in the package everything a user needs to "see what's going on" without providing their own data
#--------------------------------

# step 1: load the provided example data and parse it to the objects required for `build_xml()`
xmlGenerator::load_example_data()
# step 2: run `build_xml()` on the object produced by step 1
xmlGenerator::build_xml(record_list)
```


# An example of a "real" workflow

The only difference between the "example" workflow above and a "real" workflow is the source of object `forms_spreadsheet`. In a real workflow, the user must specify their `spreadsheet_path` and provide `spreadsheet_path` as an argument to `load_data()`.  

```{r}
#--------------------------------
# A "real" workflow
# Purpose: mock up the exact workflow a user needs to follow to use library(xmlGenerator) to make xml
#--------------------------------
# step 1: assign the filepath to your forms data as a character string
FORMS_SPREADSHEET <- "data-raw/testdata.xlsx"
# step 2: run loadData()
# this function validates your input data for filetype, colnames, reference types
# then it parses your data into a list (an instance of class record_list)
xmlGenerator::load_data(spreadsheet_path = FORMS_SPREADSHEET)
# step 3: run buildXML()
# this function builds xml out of your `record_list` object
# "write = TRUE" will bring up a dialog for the user to choose where to save the xml output
xmlGenerator::build_xml(record_list, write = FALSE)
```

# Example code

[main.R](https://github.com/cw-data/xmlGenerator/blob/main/main.R) is provides valid syntax to use `xmlGenerator`.
